---
title: "Diabetes, Neural Network"
author:"Yundi Qiu"
output: html_document
date: "2022-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Cleaning the Global Environment
```{r}
rm(list=ls(all=T))
```


Import necessary Packages and read data
```{r}
library(tidyverse)
library(patchwork)
library(caret)
library(cowplot)
library(readr)
library(neuralnet)
library(gains)

diabetes <- read_csv("diabetes.csv")
summary(diabetes)
```


Set up training and validation data
```{r}
set.seed(1234)  
train.index <- sample(c(1:dim(diabetes)[1]), dim(diabetes)[1]*0.6)
train <- diabetes[train.index, ]
valid <- diabetes[-train.index, ]
```



Normalization
method = c("scale","center")
```{r}
preprocVal <- preProcess(train[,-c(9)], method = c("center","scale"))
trainTran <- predict(preprocVal, train)
validTran <- predict(preprocVal, valid)
summary(trainTran)
```


Neural Network Model 1
hidden layers = 2, node = c(4, 2)
```{r}
nn.diabetes <- neuralnet(Outcome ~ 
                  Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin
                 + BMI + DiabetesPedigreeFunction + Age, data = trainTran, 
                 hidden = c(4, 2))

nn.diabetes$result.matrix
plot(nn.diabetes)
```


Performance Analysis
Confusion Matrix for Neural Network Model 1
```{r}
valid$predict <- predict(nn.diabetes, validTran, type="response")
train$predict <- predict(nn.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```


Lift Chart for Neural Network Model 1
```{r}
nn.diabetes.gain <- gains(validTran$Outcome, valid$predict)
nn.diabetes.gain

# plot lift chart
plot(c(0, nn.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, nn.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) 
y_baseline <- c(0, sum(validTran$Outcome)) 
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- nn.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = nn.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Neural Network Model 2
hidden layers = 3, node = c(6, 4, 2)
```{r}
nn.diabetes <- neuralnet(Outcome ~ 
                  Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin
                 + BMI + DiabetesPedigreeFunction + Age, data = trainTran, 
                 threshold = 0.1, stepmax = 1e+08, hidden = c(6, 4, 2))

nn.diabetes$result.matrix
plot(nn.diabetes)
```


Performance Analysis
Confusion Matrix for Neural Network Model 2
```{r}
valid$predict <- predict(nn.diabetes, validTran, type="response")
train$predict <- predict(nn.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```



Lift Chart for Neural Network Model 2
```{r}
nn.diabetes.gain <- gains(validTran$Outcome, valid$predict)
nn.diabetes.gain

# plot lift chart
plot(c(0, nn.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, nn.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) 
y_baseline <- c(0, sum(validTran$Outcome)) 
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- nn.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = nn.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Neural Network Model 3
hidden layers = 2, node = c(5, 2)
```{r}
nn.diabetes <- neuralnet(Outcome ~ 
                  Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin
                 + BMI + DiabetesPedigreeFunction + Age, data = trainTran, 
                hidden = c(5, 2))

nn.diabetes$result.matrix
plot(nn.diabetes)
```


Performance Analysis
Confusion Matrix for Neural Network Model 3
```{r}
valid$predict <- predict(nn.diabetes, validTran, type="response")
train$predict <- predict(nn.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```



Lift Chart for Neural Network Model 3
```{r}
nn.diabetes.gain <- gains(validTran$Outcome, valid$predict)
nn.diabetes.gain

# plot lift chart
plot(c(0, nn.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, nn.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) 
y_baseline <- c(0, sum(validTran$Outcome)) 
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- nn.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = nn.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Neural Network Model 5
(named model 5 because it is analyzed after applying "range" method in the project report)
hidden layers = 1, node = 1
```{r}
nn.diabetes <- neuralnet(Outcome ~ 
                  Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin
                 + BMI + DiabetesPedigreeFunction + Age, data = trainTran, 
                 hidden = 1)

nn.diabetes$result.matrix
plot(nn.diabetes)
```


Performance Analysis
Confusion Matrix for Neural Network Model 5
```{r}
valid$predict <- predict(nn.diabetes, validTran, type="response")
train$predict <- predict(nn.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```



Lift Chart for Neural Network Model 5
```{r}
nn.diabetes.gain <- gains(validTran$Outcome, valid$predict)
nn.diabetes.gain

# plot lift chart
plot(c(0, nn.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, nn.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) 
y_baseline <- c(0, sum(validTran$Outcome)) 
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- nn.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = nn.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```




Neural Network Model 6
(named model 6 because it is analyzed after the team tried 1 hidden layer, 1 node with "scale" and "center" normalization method)
hidden layers = 2, node = c(1, 1)
```{r}
nn.diabetes <- neuralnet(Outcome ~ 
                  Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin
                 + BMI + DiabetesPedigreeFunction + Age, data = trainTran, 
                 hidden = c(1, 1))

nn.diabetes$result.matrix
plot(nn.diabetes)
```


Performance Analysis
Confusion Matrix for Neural Network Model 6
```{r}
valid$predict <- predict(nn.diabetes, validTran, type="response")
train$predict <- predict(nn.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```



Lift Chart for Neural Network Model 6
```{r}
nn.diabetes.gain <- gains(validTran$Outcome, valid$predict)
nn.diabetes.gain

# plot lift chart
plot(c(0, nn.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, nn.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) 
y_baseline <- c(0, sum(validTran$Outcome)) 
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- nn.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = nn.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Normalization 2
method = "range"
```{r}
preprocVal <- preProcess(train[,-c(9)], method = "range")
trainTran <- predict(preprocVal, train)
validTran <- predict(preprocVal, valid)
summary(trainTran)
```



Neural Network Model 4
hidden layers = 2, node = c(6, 4, 2)
```{r}
nn.diabetes <- neuralnet(Outcome ~ 
                  Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin
                 + BMI + DiabetesPedigreeFunction + Age, data = trainTran, 
                 threshold = 0.1, stepmax = 1e+08, hidden = c(5, 2))

nn.diabetes$result.matrix
plot(nn.diabetes)
```


Performance Analysis
Confusion Matrix for Neural Network Model 4
```{r}
valid$predict <- predict(nn.diabetes, validTran, type="response")
train$predict <- predict(nn.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```



Lift Chart for Neural Network Model 4
```{r}
nn.diabetes.gain <- gains(validTran$Outcome, valid$predict)
nn.diabetes.gain

# plot lift chart
plot(c(0, nn.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, nn.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) 
y_baseline <- c(0, sum(validTran$Outcome)) 
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- nn.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = nn.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```




