---
title: "Diabetes Logistic Regression"
author:"Yundi Qiu"
output: html_document
date: "2022-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Cleaning the Global Environment
```{r}
rm(list=ls(all=T))
```


Import necessary Packages and read data
```{r}
library(tidyverse)
library(patchwork)
library(caret)
library(cowplot)
library(readr)
library(neuralnet)
library(gains)

diabetes <- read_csv("diabetes.csv")
summary(diabetes)
```


Set up training and validation data
```{r}
set.seed(1234)  
train.index <- sample(c(1:dim(diabetes)[1]), dim(diabetes)[1]*0.6)
train <- diabetes[train.index, ]
valid <- diabetes[-train.index, ]
summary(train)
```


Normalization 1
method = c("scale","center")
```{r}
preprocVal <- preProcess(train[,-c(9)], method = c("scale","center"))
trainTran <- predict(preprocVal, train)
validTran <- predict(preprocVal, valid)
```


Logistic Regression Model 1
```{r}
logit.diabetes <- glm(Outcome ~ ., data = trainTran, family="binomial")
summary(logit.diabetes)
```


Performance Analysis:
Confusion Matrix for Logistic Regression Model 1
```{r}
valid$predict <- predict(logit.diabetes, validTran, type="response")
train$predict <- predict(logit.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
``` 



Lift Chart for Logistic Regression Model 1
```{r}
logit.diabetes.gain <- gains(validTran$Outcome, valid$predict)
logit.diabetes.gain

# plot lift chart
plot(c(0, logit.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, logit.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) # length of the validation set
y_baseline <- c(0, sum(validTran$Outcome)) # The total number of Class 1 (actual) in the validation set
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- logit.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = logit.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Logistic Regression Model 2
```{r}
logit.diabetes <- glm(Outcome ~ Pregnancies + Glucose + BMI + DiabetesPedigreeFunction, data = trainTran)
summary(logit.diabetes)
```

Performance Analysis:
Confusion Matrix for Logistic Regression Model 2
```{r}
valid$predict <- predict(logit.diabetes, validTran, type="response")
train$predict <- predict(logit.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
``` 



Lift Chart for Logistic Regression Model 2
```{r}
logit.diabetes.gain <- gains(validTran$Outcome, valid$predict)
logit.diabetes.gain

# plot lift chart
plot(c(0, logit.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, logit.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) # length of the validation set
y_baseline <- c(0, sum(validTran$Outcome)) # The total number of Class 1 (actual) in the validation set
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- logit.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = logit.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```


Normalization 2
method = "range"
```{r}
preprocVal <- preProcess(train[,-c(9)], method = "range")
trainTran <- predict(preprocVal, train)
validTran <- predict(preprocVal, valid)
```


Logistic Regression Model 3
```{r}
logit.diabetes <- glm(Outcome ~ ., data = trainTran, family="binomial")
summary(logit.diabetes)
```


Performance Analysis:
Confusion Matrix for Logistic Regression Model 3
```{r}
valid$predict <- predict(logit.diabetes, validTran, type="response")
train$predict <- predict(logit.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
``` 



Lift Chart for Logistic Regression Model 3
```{r}
logit.diabetes.gain <- gains(validTran$Outcome, valid$predict)
logit.diabetes.gain

# plot lift chart
plot(c(0, logit.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, logit.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) # length of the validation set
y_baseline <- c(0, sum(validTran$Outcome)) # The total number of Class 1 (actual) in the validation set
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- logit.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = logit.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Logistic Regression Model 4
```{r}
logit.diabetes <- glm(Outcome ~ Pregnancies + Glucose + BMI + DiabetesPedigreeFunction, data = trainTran)
summary(logit.diabetes)
```


Performance Analysis:
Confusion Matrix for Logistic Regression Model 4
```{r}
valid$predict <- predict(logit.diabetes, validTran, type="response")
train$predict <- predict(logit.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
``` 



Lift Chart for Logistic Regression Model 4
```{r}
logit.diabetes.gain <- gains(validTran$Outcome, valid$predict)
logit.diabetes.gain

# plot lift chart
plot(c(0, logit.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, logit.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) # length of the validation set
y_baseline <- c(0, sum(validTran$Outcome)) # The total number of Class 1 (actual) in the validation set
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- logit.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = logit.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```



Normalization 3
method = "pca"
```{r}
preprocVal <- preProcess(train[,-c(9)], method = "pca")
trainTran <- predict(preprocVal, train)
validTran <- predict(preprocVal, valid)
```


Logistic Regression Model 5
```{r}
logit.diabetes <- glm(Outcome ~ PC1 + PC2 + PC3 + PC5 + PC6, data = trainTran)
summary(logit.diabetes)
```


Performance Analysis:
Confusion Matrix for Logistic Regression Model 5
```{r}
valid$predict <- predict(logit.diabetes, validTran, type="response")
train$predict <- predict(logit.diabetes, trainTran, type="response")

ggplot(train, aes(predict, color=as.factor(Outcome))) + 
  geom_density(size=1) + 
  ggtitle("Training Set's Predicted Score") 

cmatrix <- confusionMatrix(as.factor(ifelse(validTran$Outcome > 0.5, 1, 0)), 
                           as.factor(ifelse(valid$predict > 0.5, 1, 0)))
cmatrix
```



Lift Chart for Logistic Regression Model 5
```{r}
logit.diabetes.gain <- gains(validTran$Outcome, valid$predict)
logit.diabetes.gain

# plot lift chart
plot(c(0, logit.diabetes.gain$cume.pct.of.total*sum(validTran$Outcome)) ~ 
       c(0, logit.diabetes.gain$cume.obs), 
     xlab="# cases", ylab="Cumulative", type="l",xaxp = c(0, 2000, 10))

# Add baseline model:
x_baseline <- c(0, dim(validTran)[1]) # length of the validation set
y_baseline <- c(0, sum(validTran$Outcome)) # The total number of Class 1 (actual) in the validation set
lines(x_baseline, y_baseline, lty=2)

#compute deciles and plot decile-wise chart
heights <- logit.diabetes.gain$mean.resp/mean(validTran$Outcome)
midpoints <- barplot(heights, names.arg = logit.diabetes.gain$depth, ylim = c(0,9), 
                     xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```
